---
title: "Selecting 1000 DE genes"
output: html_document
---

#
#
##### Set working directory and load packages and data
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
setwd("PATH to your working direction")
library(edgeR)
library(DESeq)
library(DESeq2)
load("Skin.Expression.data.RData")
counts<-as.matrix(GTEX.Skin.Expr)

#or for Muscle data:
#load("Muscle.Expression.data.RData")
#counts<-as.matrix(GTEX.Muscle.Expr)
```

#
##### Filtering Genes out with less than 1 cpm in at least half of the samples 
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
d<-as.numeric(dim(counts)[2])/2
counts<-counts[which(rowSums(cpm(counts)>1) >= d),]
```

#
##### Alalyse genes for differential expression using edgeR, FDR=0.05
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
d <- DGEList(counts, group= group)
d <- calcNormFactors(d)
d <- estimateCommonDisp(d)
d <- estimateTagwiseDisp(d)
de<-exactTest(d)
DE<-de$table
padj<-p.adjust(DE$PValue, method="fdr")
DE<-cbind(DE,padj)
deSig <- DE[which(DE$padj<0.05),]
```

#
##### Alalyse genes for differential expression using DESeq2,FDR=0.05
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
pheno<-as.data.frame(group)
row.names(pheno)<- colnames(counts)
colnames(pheno)<- "condition"
ds2 <- DESeqDataSetFromMatrix(countData=counts, colData=pheno, design = ~ condition )
ds2 <- DESeq(ds2)
res2 <-results(ds2)
resSig <- res2[which(res2$padj<0.05),]
```

#
##### Alalyse genes for differential expression using DESeq,FDR=0.05
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
ds <- newCountDataSet(counts, group )
ds <- estimateSizeFactors( ds )
ds <- estimateDispersions( ds )
res <- nbinomTest(ds, "0", "1")
DeSig <- res[which(res$padj<0.05),]
```

#
##### Identify the intersection of DE genes with a predefined fold change in all three packages
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
# Set fold change (1.5, 2 or 2.5)
fc<-2

##edgeR##
deSigFC1<-deSig[which(abs(deSig$logFC) >= log2(fc)), ]
sG.names1<-as.matrix(row.names(deSigFC1))

##DESeq2##
deSigFC2<-resSig[which((abs(resSig$log2FoldChange))>= log2(fc)), ]
sG.names2<-as.matrix(row.names(deSigFC2))

##DESeq##
deSigFC3<-DeSig[which(abs(DeSig$log2FoldChange) >= log2(fc)), ]
sG.names3<-as.matrix(deSigFC3$id)

sG.names<-Reduce(intersect,list(sG.names1,sG.names2,sG.names3))
```

#
##### Randomly selecting 1000 DE genes 
```{r, echo=TRUE, results = "asis", warning = FALSE, message = FALSE, eval=FALSE} 
set.seed(679)
a<-sample(seq(1:dim(sG)[[1]]))[1:1000] 
sG.names<-sG.names[a,]
save(sG.names, counts, group, file = "selected_genes_Skin_1000_FC_2.RData" )
#save(sG.names, counts, group, file = "selected_genes_Muscle_1000_FC_2.RData" )
```
